Modeling the near-wake of a vertical-axis cross-flow turbine with 2-D and 3-D RANS

### 2016-09-25 reviews/reviewer2-response.md

# Response to reviewer 2

We thank the reviewer for his or her thoughtful comments, which have helped us
improve the paper.

>In general more work needs to be performed on
• Discussion of mesh factors that may affect the simulation results

>• Nomenclature, and the lack of definition of many of the variables will cause confusion for many readers

>• Improvement of figures with nomenclature, labels, axis etc.

>• More explanation of why the various transport properties are investigated and their applicability

>Additionally the title could be better, maybe add a Computational Fluid Dynamics or simulation somewhere?

### 2016-09-25 reviews/reviewer3-response.md

# Response to reviewer 3

We appreciate the comments given here, and addressing them has improved the manuscript significantly.

>1. For convergence study, Fig 5 gives the results for different choice of time step. however, for SST model, the Cp do not convergent with the time step. On the other hand, the choice of time step is always coupled with the choice of space step. Generally speaking, the time step cannot be too small for a certain space discretion condition. Therefore, It is suggested that the author provide this relationship between space and time discretion.

>2. No the convergence study for 3d model in this paper.

>3. In Performance Prediction session, the author gives a figure(Fig.6 ) to show the comparison of 3d , 2d and experimental data. But fig.6 only gives the results for one case for different methods. The 2d method seems over-predict the Cp and Cd. more results of different cases should be given to verify this idea.

>4. Shaft and Arm/Strut effects are important to the turbine, and have been discussed by previous publications by using RANS such as
>
Behrouzi F, Malik A M A, Nakisa M, et al. Arm Effect on Performance of Vertical Axis Current Turbine Using RANS Simulation. Applied Mechanics & Materials, 2014, 554(554):531-535.
>
and by using experimental test such as
Li Y, Calisal S M. Three-dimensional effects and arm effects on modeling a vertical axis tidal current turbine. Renewable Energy, 2010, 35(10):2325-2334.
>
However, the authors have not well explained it in the manuscript. Please extend this part.

### 2016-09-25 reviews/reviewer3-response.md

# Response to reviewer 3

We appreciate the comments given here, and addressing them has improved the manuscript significantly.

>1. For convergence study, Fig 5 gives the results for different choice of time step. however, for SST model, the Cp do not convergent with the time step. On the other hand, the choice of time step is always coupled with the choice of space step. Generally speaking, the time step cannot be too small for a certain space discretion condition. Therefore, It is suggested that the author provide this relationship between space and time discretion.

It is true that in general spatial and temporal discretization are coupled.
However, using the PIMPLE solver in OpenFOAM allows for stable solutions even
when the CFL condition is not satisfied in every cell. In our case, we used a
fixed time step, such that it would not be changed by the solver based on the
Courant number. For example, the temporal convergence data shown were acquired
using the "final" spatial resolution and vice versa. This did take some
iterative process, but ultimately we have one independent variable for each
plot. A note has been added to the manuscript to clarify this point.

### 2016-09-25 paper.tex
Clarify wording on BC description
Clarify convergence study strategy

Initial and boundary conditions were set to match those of the tow tank as well
as possible. The velocity at the inlet, bottom, and side walls was fixed to 1
m/s to match the tow tank case, while the top boundary condition was a slip
velocity condition. Pressure was held fixed at the outlet while a zero-gradient
boundary condition was applied at the inlet (a typical
velocity-inlet/pressure-outlet case). Note that in 2-D the top and bottom
boundary conditions are ``empty,'' which is an OpenFOAM convention to indicate
two-dimensionality.

@@

Time step dependence was evaluated using the 2-D $N_x=70$ grid, the results from
which are shown in Figure~\ref{fig:2d-br-verification}. It was seen that the
Spalart--Allmaras model converged well with decreasing time step, leading to a
final time step of 0.001 s. The results from the SST model show a local minimum
at $\Delta t = 0.002$ s, with some divergence for smaller time steps. The local
minimum was chosen as the final time step to run the simulations. Note that the
SST model's convergence behavior may be due to its specific implementation in
OpenFOAM, and not indicative of the nature of the model equations. Verification
studies for CFTs with this level of detail in the literature are not common,
though the final time step is comparable to others \cite{Balduzzi2016}.

>>>>

Initial and boundary conditions were set to approximate those of the tow tank as
closely as possible. The velocity at the inlet, bottom, and side walls was fixed
to 1 m/s to match the tow tank case, while the top boundary condition was a slip
velocity condition. Pressure was held fixed at the outlet while a zero-gradient
boundary condition was applied at the inlet (a typical
velocity-inlet/pressure-outlet case). Note that in 2-D the top and bottom
boundary conditions are ``empty,'' which is an OpenFOAM convention to indicate
two-dimensionality.

@@

Time step dependence was evaluated using the 2-D $N_x=70$ grid, the results from
which are shown in Figure~\ref{fig:2d-br-verification}. It was seen that the
Spalart--Allmaras model converged well with decreasing time step, leading to a
final time step of 0.001 s. The results from the SST model show a local minimum
at $\Delta t = 0.002$ s, with some divergence for smaller time steps. The local
minimum was chosen as the final time step to run the simulations. Note that the
SST model's convergence behavior may be due to its specific implementation in
OpenFOAM, and not indicative of the nature of the model equations. Verification
studies for CFTs with this level of detail in the literature are not common,
though the final time step is comparable to others \cite{Balduzzi2016}. Note
that while in general there may be a coupling between necessary spatial and
temporal resolution, these values were arrived at iteratively, i.e., the spatial
dependence was varied with the chosen time step, and vice versa, such that there
was one independent variable for each parameter sweep.
